Effectful Semantics for Substructural SSA

### 2023-11-16 pldi24/paper.tex
foo

\newcommand{\todo}[1]{\stepcounter{todos} \textcolor{red}{TODO \arabic{todos}:} #1}

@@

Now that we've given a semantics to \isotopessa programs, we can begin to use it
to develop an equational theory for performing optimizations. Since we're
approaching things from a type-theoretic perspective, we begin by addressing the
semantic versions of the standard syntactic metatheorems of weakening and
substitution.

\todo{Move weakening and upgrade lemmas to appendix}

% In particular, semantic weakening can be stated in the expected
% manner: if we have an expression or block which is well typed in both \(\Gamma\)
% and \(\Delta\), where \(\cwk{\Gamma}{\Delta}\), then evaluating it in \(\Gamma\)
% is the same as first strengthening \(\Gamma\) to \(\Delta\), dropping unused
% affine variables, and then evaluating it in \(\Delta\).
% \begin{lemma}[Semantic Weakening] 
%   If \(\cwk{\Gamma}{\Delta}\), then 
%   \begin{itemize}
%     \item For all \(\hasty{\Delta}{p}{a}{A}{q}\),
%     \[
%       \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
%       ; \dnt{\hasty{\Delta}{p}{a}{A}{q}}
%       = \dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
%     \]
%     \item For all \(\haslb{\Delta}{t}{\ms{L}}\),
%     \[
%       \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
%       ; \dnt{\haslb{\Delta}{t}{\ms{L}}}
%       = \dnt{\haslb{\Gamma}{t}{\ms{L}}}  
%     \]
%   \end{itemize}
% 
%   Similarly, if \(\lwk{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), then
%   \[
%     \dnt{\haslb{\Gamma}{t}{\ms{L}}}
%     ; \dnt{\lwk{\ms{L}}{\ms{K}}}
%     = \dnt{\haslb{\Gamma}{t}{\ms{K}}}
%   \]
% \end{lemma}
% 
% \TODO{some text about upgrade}
% 
% \begin{lemma}[Semantic Upgrade]
%   For all \(\Gamma, a, A\), if \(\hasty{\Gamma}{1}{a}{A}{q}\), then
%   \[\dnt{\hasty{\Gamma}{0}{a}{A}{q}} = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{}\]
% \end{lemma}

@@

\TODO{some text about substitution semantics}

\TODO{some text about semantic substitution}

>>>>

\newcommand{\todo}[1]{\stepcounter{todos} \textcolor{red}{TODO \arabic{todos}: #1}}

@@

Now that we have a semantics for \isotopessa programs, we can show
that it matches the syntax by showing that the syntactic and semantic
notions of substitution coincide.

\todo{Move weakening and upgrade lemmas to appendix}
% Since we're
% approaching things from a type-theoretic perspective, we begin by addressing the
% semantic versions of the standard syntactic metatheorems of weakening and
% substitution. 
% 
% In particular, semantic weakening can be stated in the expected
% manner: if we have an expression or block which is well typed in both \(\Gamma\)
% and \(\Delta\), where \(\cwk{\Gamma}{\Delta}\), then evaluating it in \(\Gamma\)
% is the same as first strengthening \(\Gamma\) to \(\Delta\), dropping unused
% affine variables, and then evaluating it in \(\Delta\).
% \begin{lemma}[Semantic Weakening] 
%   If \(\cwk{\Gamma}{\Delta}\), then 
%   \begin{itemize}
%     \item For all \(\hasty{\Delta}{p}{a}{A}{q}\),
%     \[
%       \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
%       ; \dnt{\hasty{\Delta}{p}{a}{A}{q}}
%       = \dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
%     \]
%     \item For all \(\haslb{\Delta}{t}{\ms{L}}\),
%     \[
%       \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
%       ; \dnt{\haslb{\Delta}{t}{\ms{L}}}
%       = \dnt{\haslb{\Gamma}{t}{\ms{L}}}  
%     \]
%   \end{itemize}
% 
%   Similarly, if \(\lwk{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), then
%   \[
%     \dnt{\haslb{\Gamma}{t}{\ms{L}}}
%     ; \dnt{\lwk{\ms{L}}{\ms{K}}}
%     = \dnt{\haslb{\Gamma}{t}{\ms{K}}}
%   \]
% \end{lemma}
% 
% \TODO{some text about upgrade}
% 
% \begin{lemma}[Semantic Upgrade]
%   For all \(\Gamma, a, A\), if \(\hasty{\Gamma}{1}{a}{A}{q}\), then
%   \[\dnt{\hasty{\Gamma}{0}{a}{A}{q}} = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{}\]
% \end{lemma}

@@

\TODO{some text about substitution semantics -- substitution = composition of env semantics with therm semantics}
\TODO{some text about semantic substitution -- it's a function that does the obvious thing}
